\myTitle

### 2024-01-22 body-text-of-paper.tex
small tweaks
Add an annotation
More things to the examples methods and results.

@@

NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE SHOULD NOTE
HOW THOSE CAME OUT.

@@

\subsection*{Empirical Examples}

We illustrate the use of \gscramble{} in two examples.  The first involves
simulating a small number ($<\,$70) of markers that are fixed, or nearly
fixed, for alternate alleles between two species of trout: rainbow trout ({\em 
Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
simulation study is to determine how accurately individuals of different hybrid
categories between the species can be identified, and whether the use of linked 
markers deflates the estimates of uncertainty to an unacceptable degree.  The 
second example involves many more markers ($\approx$14,000) and 
investigates the accuracy for
identifying individuals admixed between closely related groups of feral
pigs in Missouri.

@@

Figure is completed and is in Fig.~\ref{fig:ad-fract-sims}.
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
\subsection*{Empirical Examples}

@@

Figure is completed and is Fig.~\ref{fig:newhyb-sims-barplot}
%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated an additional  35 markers, some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%

>>>>

\subsubsection*{RISPI when estimating admixture fractions}
\label{sec:rispi-sim}

@@

%NOTE, WE ALSO DID THESE SIMULATIONS USING GSCRAMBLE AND WE COULD NOTE
%HOW THOSE CAME OUT.  
% BUT MAYBE JUST HAVE THAT IN CASE A REFEREE ASKS ABOUT IT.

@@

% I THOUGHT ABOUT PUTTING THESE RESULTS IN, BUT I
% THINK WE ARE FINE WITHOUT THEM.
%\subsection{Simulation: \gscramble{} Eliminates RISPI}
%
%We present a brief simulation similar to those in
%section ``RISPI when estimating admixture fractions,'' but conducted using
%\gscramble{} rather than sampling with replacement.  We focused on two
%scenarios that

\subsection*{Examples}

We illustrate the use of \gscramble{} in three example cases. The first is
a simple simulatino to show that, unlike sampling with replacement, the
approach of \gscramble{} does not produce RISPI.
The next two are examples using empirical data sets.
The first of these involves
simulating a small number ($<\,$70) of markers that are fixed, or nearly
fixed, for alternate alleles between two species of trout: rainbow trout ({\em 
Oncorhynchus mykiss}) and cutthroat trout ({\em O. clarkii}). The goal of the 
simulation study is to determine how accurately individuals of different hybrid
categories between the species can be identified, and whether the use of linked 
markers deflates the estimates of uncertainty to an unacceptable degree.  The 
second example involves many more markers ($\approx$14,000) and 
investigates the accuracy for
identifying individuals admixed between closely related groups of feral
pigs in Missouri.

\subsubsection*{No RISPI when using \gscramble{}}

For this example, we repeat the simulation described in \nameref{sec:rispi-sim};
however, after simulating individuals in the two samples drawn from the same population,
we create data sets of admixed individuals using \gscramble{}, rather than sampling
gene copies with replacement.  As before, individuals were simulated with fractions of
ancestry from sample $A$ in $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$, but these were simulated from pedigrees using gscramble.  We limited
this simulation to one of the scenarios that showed extreme RISPI when sampling gene
copies with replacement: $L=1,000$ markers with sizes of samples $A$ and $B$ being
$N=50$. This simulation was repeated 240 times, and the results of each analyzed with
ADMIXTURE to estimate the ancestry fractions of each simulated individual.

@@

When hybrids were simulated in a way that accounts for the physical linkage of
markers, the variance in admixture fractions was considerably higher than when
markers were simulated under the assumption that none of them are physically
linked. (Fig.~\ref{fig:ad-fract-sims}).
%%%%%%%%%%
\begin{figure}
\newcommand{\adfcap}{\footnotesize Admixture fractions of individuals simulated
assuming no physical linkage (top four panels)
and when physical linkage is simulated using \gscramble{}.  In the top 4 panels the number
of simulated loci appears in the panel header.  Failing to simulate markers with physical linkage
produces distributions of admixture fractions that are far more narrow than reality, giving the
mistaken impression that admixture fractions of different hybrid categories do not overlap.  }
\includegraphics[width=\columnwidth]{figures/link-and-nolink-sims-crop.pdf}
\caption[\adfcap]{\adfcap}
\label{fig:ad-fract-sims}
\end{figure}
%%%%%%%%%%
If markers are assumed to be unlinked, then the variance around the $Q$ estimates of different 
hybrid categories continually declines as more markers are added, from 100, to 100,000, which 
makes it appear that different hybrid categories become increasingly well-resolved.  For example, 
looking at the distributions for 100,000 unlinked markers one would think that it would be easy to 
distinguish between the $BC_1$ through $BC_4$ categories on the basis of the admixture fractions.
This is, however, not the case.  The bottom panel of Fig.\ref{fig:ad-fract-sims} shows the distribution
of the true admixture fraction (calculated as the total length of genome from each of the two
different species) of each individual when simulated with physical linkage using \gscramble{}.
When physical linkage is not ignored, it is quite clear that there is considerable overlap in the
admixture fractions between the $BC_1$ and $BC_2$ categories, and that $BC_2$, $BC_3$, and 
$BC_4$ all have admixture fraction distributions that overlap greatly.  This shows that,
in real life, when physical linkage cannot be ignored, it may be quite difficult to distinguish
between different backcross categories on the basis of their admixture fractions.

\subsection*{Examples}

\subsubsection*{No RISPI when using \gscramble{}}

As expected, there was no RISPI apparent in the ADMIXTURE-estimated
ancestry proportions  when simulating individuals
using \gscramble{} (Fig.~\ref{fig:norispi}).  On average, the estimated $Q$ values
were around 0.5, which is expected when the samples are drawn from the same
population.
%%%
\begin{figure}
\newcommand{\gssimcap}{\footnotesize ADMIXTURE estimates of ancestry
fractions, $Q$, for individuals simulated from samples from the same population,
simulated using \gscramble{}.  There is a considerable variation in the estimates,
but the central tendency of them all is around 0.5, as would be expected in this case
within which there is not genetic basis for distinguishing individuals from the different
samples.  Contrast this to the corresponding panel ($L=1,000, N=50$) in Fig.~\ref{fig:bias-sims}(b),
in which the RISPI is readily apparent.}
\includegraphics[width=\columnwidth]{figures/gscram_supervised_Qs.pdf}
\caption[\gssimcap]{\gssimcap}
\label{fig:norispi}
\end{figure}
%%%

@@

The barplots in Fig.~\ref{fig:newhyb-sims-barplot} show the distribution of posterior
probabilities of the different hybrid categories.

%%%%%%%%%%
\begin{figure}
\newcommand{\nhbpcap}{\footnotesize Posterior probabilities for simulated
individuals of different trout hybrid categories.  Each pair of plots shows the results
for individuals of the true hybrid category listed atop the pair.  (See Table~\protect\ref{tab:newhybcats} for meaning of the names.)  The top panel of each pair
shows results for fish simulated with the 30 markers used by \protect\citet{rizza2023limited}, and the bottom panel shows results for fish simulated with the original 30 markers,
plus an additional  35 (65 total), some of which are linked
on the same chromosome. Each individual appears as a thin vertical line, posterior probabilities of
each inferred hybrid category are indicated by extent of different colors on those lines.  Individuals
are sorted according to maximum {\em a posteriori} hybrid category, and, within that by the
maximum posterior probability. }
{\centering
\includegraphics[width=0.98\columnwidth]{figures/newhybs-texed-crop.pdf}
}
\caption[\nhbpcap]{\nhbpcap}
\label{fig:newhyb-sims-barplot}
\end{figure}
%%%%%%%%%%
These are similar to ADMIXTURE barplots;
however, rather than plotting the ancestry fraction, the different colors are showing the
posterior probabilities of different hybrid categories.  Each individual is represented by a thin
vertical line, and the individuals have been arranged from left to right and the colors have been
arranged from top to bottom throughout different sections of the plot to make it easy to see
the patterns.  The {\em maximum-a-posteriori} (MAP) hybrid category for any individual is represented
by the color that is on the bottom of the bar, so it is easy to see the shifts where the MAP hybrid
category changes from one to another.  It is apparent over all individuals simulated to be
either CCT, $F_1$, $F_2$, $BC_1$--CCT, or $BC_2$--CCT, that the MAP estimate of their
hybrid categories were correct for largely the same number of individuals whether 30 or 65 markers
were used. By contrast, for the individuals simulated to be of the complex hybrid category
$BC_1$--CCT~$\times$~$BC_2$--CCT, the MAP estimate from 65 markers was correct more often 
than from 
30 markers.  However, it is also apparent that using 65 markers, some of which are physically
linked, in the software {\em NewHybrids}---which assumes that all markers are unlinked---yields 
estimates of uncertainty that are inaccurate.  Tellingly, when the estimates using 65 markers are 
incorrect (for example, when $BC_1$--CCT individuals have a maximum posterior of being in the 
$BC_2$--CCT category,
or when individuals that are truly of the $BC_1$--CCT~$\times$~$BC_2$--CCT category have a
maximum posterior of being from the $BC_2$--CCT category), many of those incorrect posteriors
are quite high ($>0.90$).  In fact, using 65 markers yields more incorrect {\em and high} posteriors 
than does using the 30 unlinked markers. So, as expected, ignoring physical linkage the way
{\sc NewHybrids} does can yield a poor estimate of the certainty around different inferences.
